\emph{Light Exposure Behavior Assessment (LEBA)}: Development of a novel instrument to capture light exposure-related behaviours

### 2022-01-02 appendix-1.Rmd
Sup.files updated

```{r factormp, warning=F}
map <- readr::read_csv("Table_raw/map_stat")
map <- map[,-1]
papaja::apa_table(map,align = "c", caption = "MAP Statistics",  font_size = "small")
```

@@

>>>>

```{r factormp, warning=F}
map <- readr::read_csv("Table_raw/map_stat")
map1 <- (map$`MAP Statistic`)

map1 <- MOTE::apa(map1,5,F)
map1 <- as.character(map1)
map1 <- as.data.frame(map1)
colnames(map1) <- "MAP Statistics"
map2 <- map[,-c(1,2)]
map3 <- cbind(map1, map2)

papaja::apa_table(map3, caption = "Map Statistics")
```

@@

```{r MI, include=F, warning=F}
data <- readRDS("leba_2021-09-08.rds")

# Separating EFA and CFA samples with descriptive column
descriptives.data <- data

## Merge "0"s and "1"s into "1"s select,subset assign)
descriptives.data[ , 9:56 ][ descriptives.data[ , 9:56 ] == 0 ] <- 1
#EFA.descriptives <- descriptives.data[1:428,]
EFA.descriptives <- subset(descriptives.data, IncludedInEFA == "TRUE")
#CFA.descriptives <- descriptives.data[429:690, ]
CFA.descriptives<- subset(descriptives.data, IncludedInEFA == "FALSE")

mi.desc_table <- CFA.descriptives[,c(1,2,5,6,7)] %>%
  tbl_summary(
    by = slypos_demographics_language.factor,
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
    digits = all_continuous() ~ 2,
    label = list(slypos_demographics_age ~ "Age",
                 slypos_demographics_sex.factor ~ "Sex",
                 slypos_demographics_work_or_school.factor ~ "Occupational Status",
                 slypos_demographics_school.factor ~ "Occupational setting"
                 ),
    missing = "no"
    ) %>% add_overall() %>% bold_labels() %>% add_p() %>% add_q() %>%
modify_header(label ~ "**Variable**") %>% separate_p_footnotes() %>%
  modify_caption("Demographic Characteristics: Native English Speakers")

mides_kable <-  as_kable_extra(mi.desc_table, format = "latex",booktabs = T) 
```

```{r midesp, warning=F, message=F}
mides_kable %>% 
  kable_styling(latex_options = c("scale_down")) 
```

### 2022-01-02 appendix-2.Rmd
Sup.files updated

```{r MinResTab}
papaja::apa_table(minresTab,  caption = "Factor loadings and communality of the retained items(Minmum Residual)", note = "Only loading higher than .30 is reported", longtable=T)
```

@@

# Residuals
residual.6F =residuals(fa.6F.3 , diag = FALSE, na.rm =T)
#Count of numbers of residuals>.05
BigRRR= sum(residual.6F>abs(.05), na.rm =T)
print(BigRRR)

@@

#hist(residual.6F, main = NULL, xlab = "Six factor Solution: Residuals",ylim=c(0, 400))
#abline(h =65, lwd = 2, lty = 2)
#abline(v =.05,lwd = 2, lty = 2)
sixFac <- tabledown::fac.tab(fa.6F.3, cut =.3,complexity = F)
```

```{r sixFacTab, results='asis'}
papaja::apa_table(sixFac,  caption = "Factor loadings and communality of the retained items(six factor)", note = "Only loading higher than .30 is reported",longtable =T)
```

# Factor Analysis with Unmerged Response Option

@@

```{r EFAassumptionsAppB, include=FALSE}

#KMO test
KMO.1 <- KMO(EFA.data)

# Test of correlation matrix
bartlet.1 <- cortest.bartlett(EFA.data, n =428)

# Univariate normality

#creating normality function
shapiro_test_df <- function(df, bonf= TRUE, alpha= 0.05) {
  l <- lapply(df, shapiro.test)
  s <- do.call("c", lapply(l, "[[", 1))
  p <- do.call("c", lapply(l, "[[", 2))
  if (bonf == TRUE) {
    sig <- ifelse(p > alpha / length(l), "H0", "*")
  } else {
    sig <- ifelse(p > alpha, "H0", "*")
  }
  return(list(statistic= s,
              p.value= p,
              significance= sig,
              method= ifelse(bonf == TRUE, "Shapiro-Wilks test with Bonferroni Correction","Shapiro-Wilks test without Bonferroni Correction")))
}

Shapiro.efa <- shapiro_test_df(EFA.data)

efa.normality.tab = matrix(nrow = 48, ncol = 2)
rownameprefix <- "Item"
rownamesufix <- seq(1:48)
my.item.names <- paste(rownameprefix,rownamesufix, sep = "" )
rownames(efa.normality.tab) <- my.item.names
colnames(efa.normality.tab) <- c("statistic", "p")

efa.normality.tab[,1] <- apa(Shapiro.efa$statistic,2,T)
efa.normality.tab[,2] <- apa(Shapiro.efa$p.value,2,T)

efa.normalityprefix <- efa.normality.tab[,1]
efa.normalitysufix <- Shapiro.efa$significance
efa.normality.stat <- paste(efa.normalityprefix,efa.normalitysufix, sep = "" )

# Multivariate Normality
mardia.1 <- mardia(EFA.data, na.rm = T, plot =F)

# Descriptive Stats
Descriptives <- psych::describe(EFA.data)
alpha <- psych::alpha(EFA.data)
Item.total <- alpha$item.stats

Des.combined <- cbind(apa(Descriptives$mean,2,T), apa(Descriptives$sd,2,T), apa(Descriptives$skew,2,T),apa(Descriptives$kurtosis,2,T), efa.normality.stat,  apa(Item.total$r.cor,2,F))

colnames(Des.combined) = c("Mean", "SD", "Skew", "Kurtosis", "Shapiro-Wilk Statistics","Item-Total Correlation")
rownameprefix <- "Item"
rownamesufix <- seq(1:48)
my.names <- paste(rownameprefix,rownamesufix, sep = "" )
rownames(Des.combined) = (my.names)
```

```{r CorrMatrixAppB, include=FALSE, warning=F,message=F}
#Polychoric Correlation matrix

correlations <- polychoric(EFA.data, correct = 0)
upper <- correlations$rho[upper.tri(correlations$rho, diag = F)]

min.cor.1 <- apa(min(abs(upper)),2,F) #minimum cor.coefficient of the matrix
max.cor.1 <- apa(max(abs(upper)),2,F) # max. correlation coefficient of the matrix

determinets <- det(correlations$rho) 
# Calculating the percentage of correlations higher than .30

BigR = sum(correlations$rho >= abs(.30) & correlations$rho < abs(1.0), na.rm =T)/2 
totR = length(EFA.data)*(length(EFA.data)-1)/2
cor.per.1 <- print (BigR/totR)*100
```

```{r ParallelEFAAppB, include=FALSE}

# Parallel analysis
Horn <- paran(correlations$rho, iterations=500, centile=0, quietly=FALSE,
           status=TRUE, all=FALSE, cfa=FALSE, graph=TRUE,
           color=TRUE, col=c("black","red","blue"),
           lty=c(1,2,3), lwd=1, legend=TRUE, 
           width=640, height=640, grdevice=png, seed=0, mat=NA, n=NA)

Adjusted.EV <- data.frame(Horn$AdjEv)
Adjusted.EV$type = c('Adjusted Ev')
Adjusted.EV$num = c(row.names(Adjusted.EV))
Adjusted.EV$num = as.numeric(Adjusted.EV$num)
colnames(Adjusted.EV) = c('eigenvalue', 'type', 'num')

Random.EV <- data.frame(Horn$RndEv)
Random.EV$type = c('Random EV')
Random.EV$num = c(row.names(Random.EV))
Random.EV$num = as.numeric(Random.EV$num)
colnames(Random.EV) = c('eigenvalue', 'type', 'num')

Unadjusted.EV <- data.frame(Horn$Ev)
Unadjusted.EV$type = c('Unadjusted EV')
Unadjusted.EV$num = c(row.names(Unadjusted.EV))
Unadjusted.EV$num = as.numeric(Unadjusted.EV$num)
colnames(Unadjusted.EV) = c('eigenvalue', 'type', 'num')

parallel = ggplot(Adjusted.EV, aes(x=num, y=eigenvalue, shape=type, color=type)) +
  #Add lines connecting data points
  geom_line()+
    geom_line(data = Random.EV) +
  geom_line(data = Unadjusted.EV)+
    #Add the data points.
  geom_point(size=1)+
  geom_point(data = Random.EV, size=1)+
  geom_point(data = Unadjusted.EV, size=1)+
  scale_y_continuous(name="Eigen Value", limits=c(0, 10), breaks=seq(0,10,2)) +
   #Label the x-axis 'Factor Number', and ensure that it ranges from 1-max # of factors, increasing by one with each 'tick' mark.
  scale_x_continuous(name='Factor Number', limits=c(0, 48),breaks=seq(0,48,4))+
    #Add vertical line indicating parallel analysis suggested max # of factors to retain
  geom_hline(yintercept=1, linetype = 'dashed') + apatheme

```{r ScreePlotAppB, include=FALSE}
Scree <- scree(correlations$rho,factors=TRUE,pc=TRUE,main="(B)",
      hline=NULL,add=F)

FA.Scree <- data.frame(Scree$fv)
FA.Scree$type = c('Factor Analysis')
FA.Scree$num = c(row.names(FA.Scree))
FA.Scree$num = as.numeric(FA.Scree$num)
colnames(FA.Scree) = c('eigenvalue', 'type', 'num')

PA.Scree <- data.frame(Scree$pcv)
PA.Scree$type = c('Principal Component Analysis')
PA.Scree$num = c(row.names(PA.Scree))
PA.Scree$num = as.numeric(PA.Scree$num)
colnames(PA.Scree) = c('eigenvalue', 'type', 'num')

scree.plot <-ggplot(FA.Scree, aes(x=num, y=eigenvalue,color=type, shape=type))+
   geom_line()+
  geom_line(data = PA.Scree)+ 
  geom_point(size=1)+
  geom_point(data = PA.Scree, size=1)+
  scale_y_continuous(name="Eigen Value", limits=c(0, 10), breaks=seq(0,10,2)) +
  scale_x_continuous(name='Factor Number', limits=c(0, 48),breaks=seq(0,48,4))+
  geom_hline(yintercept=1, linetype = 'dashed') + apatheme 
```

```{r MAPefaAppB, include=FALSE}
#MAP
psych::VSS(EFA.data, rotate = "varimax", fm = 'minres', n.obs =428 )
```

Table \@ref(tab:tabDesAppB) summarizes the univariate descriptive statistics for the 48 items with un-merged options. Some of the items were skewed with high Kurtosis values. Our data violated both univariate normality (Shapiro-Wilk statistics) and multivariate normality assumptions [Marida's test]. Multivariate skew was = `r printnum(mardia.1$b1p)` (p <0.001) and multivariate kurtosis was = `r printnum(mardia.1$b2p)` (p <0.001). Due to these violations and ordinal nature of the response data polychoric correlations over Pearson's correlations was chosen. Sampling adequacy was checked using Kaiser-Meyer-Olkin (KMO) measures of sampling adequacy. The overall KMO vale for 48 items was `r printnum(KMO.1$MSA)` which was above the cutoff value (.50) indicating a mediocre sample. Bartlett's test of sphericity, $\chi^2$ (`r bartlet.1$df`) = `r apa(bartlet.1$chisq, 2, T)`, p < .001 indicated the correlations between items are adequate for the EFA. However only `r apa(cor.per.1, 2,T)`% of the inter-item correlation coefficients were greater than .30. The absolute value of inter-item correlation ranged between `r min.cor.1` to `r max.cor.1`.Figure \@ref(fig:corunm) depicts the correlation matrix. For un-merged response option Horn's parallel analysis with 500 iterations indicated a five-factor solution. However, Scree plot and the MAP method suggested 6-factor solution. five-factor solution . As a result, we tested both five-factor and six-factor solutions. The six factor solution yielded a factor with only two salient loading (Table \@ref(tab:EFAsix). Thus we reject the six factor solution. The five factor solution retained 24 items (Table \@ref(tab:efaunmerged)). However the factors are less interpretable in terms of common theme. Thus we reject the five factor solution.

```{r tabDesAppB, results='asis'}
apa_table(Des.combined, caption = "Descriptive Statistics for Unmerged response options",  note = "*p<.001", longtable =T)
```

```{r corunm, fig.cap="Correlation plot of the items [Unmerged response options]",  out.width="100%"}

ggcorrplot(correlations$rho, hc.order = TRUE, outline.col = "white", type = "lower", 
           ggtheme = ggplot2::theme_minimal, tl.cex = 6,tl.srt = 90 )
#p.mat <- cor_to_p(correlations$rho, 428, method = "polychoric")
#to show pvalue in the fig add p.mat=p.mat
```

```{r facIdFigAppB,  fig.cap='Factor Identification  (A) Parallel analysis (B) Scree Plot [Unmerged response options]', warning=FALSE}
cowplot::plot_grid(parallel, scree.plot, labels = "AUTO",ncol=1, label_size = 8,align = "v")

@@

```{r efaunmerged, results='asis'}
apa_table(efa.table, caption = "Factor loadings and communality of the retained items in five factor solution [Unmerged Responses]", note = "Only loading higher than .30 is reported", longtable =T)
```

@@

```{r EFAsix, results='asis'}
apa_table(efa.table.2, caption = "Factor loadings and communality of the retained items in six factor solution [Unmerged Responses]", note = "Only loading higher than .30 is reported", longtable =T)
```

# Items Retained in the Five Factor Solution [Unmerged Responses]

@@

```{r EFAResidualsAppB, eval=FALSE, fig.cap=' Histogram of residulas:  five-factor solution', warning=FALSE, include=FALSE, results='asis'}
hist(residual.5F,  main = NULL, xlab = "Five-factor solution", ylim=c(0, 400))
abline(h =35, lwd = 2, lty = 2)
abline(v =.05,lwd = 2, lty = 2)
```

# Geographic Locations of Survey Participants
```{r Timetab, include=FALSE, warning=F,message=FALSE}
descriptives.data%>%
  tbl_summary(
    label = slypos_demographics_tz.factor ~ "Time zone - Country",
    statistic = list(all_categorical() ~ "{n} ({p}%)"),
    missing = "no",
    sort = all_categorical() ~ "frequency",
    include = "slypos_demographics_tz.factor"
  )  %>% bold_labels() %>% 
  modify_header(label ~ "") %>%
  modify_caption("Geographical Location") -> tz_table

>>>>

```{r MinResTab}
papaja::apa_table(minresTab,  caption = "Factor loadings and communality of the retained items (Minmum Residual)", note = "Only loading higher than .30 is reported", longtable=T, font_size = "small")
```

@@

reduced.model.6F.3 <- dplyr::select(EFA.data, -c(item34, item44, item48, item43, item39, item22, item2, item20, item21, item24,item30, item14, item28, item6, item15, item23, item18, item42, item47, item41, item5, item13, item29))

correlations.red.6F.3 <- polychoric(reduced.model.6F.3,correct = 0)

fa.6F.4 <- fa(r=correlations.red.6F.3$rho, nfactors = 6, fm= "pa",rotate ="varimax",
              residuals = TRUE, SMC = TRUE, n.obs =428, max.iter = 500)

ll <- print(fa.6F.4, cut = .3, digits = 3, sort = TRUE)

reduced.model.6F.4 <- dplyr::select(EFA.data, -c(item34, item44, item48, item43, item39, item22, item2, item20, item21, item24,item30, item14, item28, item6, item15, item23, item18, item42, item47, item41, item5, item13, item29, item19))

correlations.red.6F.4 <- polychoric(reduced.model.6F.4,correct = 0)

fa.6F.5 <- fa(r=correlations.red.6F.4$rho, nfactors = 6, fm= "pa",rotate ="varimax",
              residuals = TRUE, SMC = TRUE, n.obs =428, max.iter = 500)

mm <- print(fa.6F.5, cut = .3, digits = 3, sort = TRUE)

# Residuals
residual.6F =residuals(fa.6F.5 , diag = FALSE, na.rm =T)
#Count of numbers of residuals>.05
BigRRR= sum(residual.6F>abs(.05), na.rm =T)
print(BigRRR)

@@

#hist(residual.6F, main = NULL, xlab = "Six factor Solution: Residuals",ylim=c(0, 400))
#abline(h =65, lwd = 2, lty = 2)
#abline(v =.05,lwd = 2, lty = 2)
sixFac <- tabledown::fac.tab(fa.6F.5, cut =.3,complexity = F)
```

```{r sixFacTab, results='asis'}
papaja::apa_table(sixFac,  caption = "Factor loadings and communality of the retained items(six factor)", note = "Only loading higher than .30 is reported; Sixth factor has only two salient loadings",longtable =T, font_size = "small")
```

@@

@@

```{r efaunmerged, results='asis'}
apa_table(efa.table, caption = "Factor loadings and communality of the retained items in five factor solution [Unmerged Responses]", note = "Only loading higher than .30 is reported", longtable =T, font_size = "small")
```

@@

```{r EFAsix, results='asis'}
apa_table(efa.table.2, caption = "Factor loadings and communality of the retained items in six factor solution [Unmerged Responses]", note = "Only loading higher than .30 is reported", longtable =T, font_size = "small")
```

## Items Retained in the Five Factor Solution [Unmerged Responses]

@@

```{r Timetab, include=FALSE, warning=F,message=FALSE}
descriptives.data%>%
  tbl_summary(
    label = slypos_demographics_tz.factor ~ "Time zone - Country",
    statistic = list(all_categorical() ~ "{n} ({p}%)"),
    missing = "no",
    sort = all_categorical() ~ "frequency",
    include = "slypos_demographics_tz.factor"
  )  %>% bold_labels() %>% 
  modify_header(label ~ "") %>%
  modify_caption("Geographical Location") -> tz_table
